---
title: "Fundamentals statistics EXAM: insurance"
author: "mathieu Bernardi"
date: "15 juillet 2019"
---

To make their own profits, the insurance company(insurer) must collect more premiums than the amount paid to the insured person.

Try to create a model to predict health care cost?


Understanding the data:

Age: insurance contractor age, years

Sex: insurance contractor gender, [female, male]

BMI: Body mass index, providing an understanding of body, weights that are relatively high or low relative to height, objective index of body weight (kg / m ^ 2) using the ratio of height to weight, ideally 18.5 to 24.9

Children: number of children covered by health insurance / Number of dependents

Smoker: smoking, [yes, no]

Region: the beneficiary’s residential area in the US, [northeast, southeast, southwest, northwest]

Charges: Individual medical costs billed by health insurance, $ #predicted value

```{r}
# Load libraries
library(ggplot2)
library(ggthemes)
library(psych)
library(dplyr)

```

```{r}
# Set the path
setwd('D:/STATS/exam/')
# Read in our dataset
assurance=read.csv("insurance.csv")

#look at the dataset
head(assurance)

```

```{r}
str(assurance)
```

```{r}
# Descriptive Statistics
summary(assurance)

```
As we can see, we are working with a small dataset with only 1338 observations and 7 variables. We would try to predict the charges variable.

```{r}
# describe assurance by region
describeBy(assurance$charges,assurance$region)
```
```{r}
ggplot(data = assurance,aes(region,charges)) + geom_boxplot(fill = c(2:5)) +
  theme_classic() + ggtitle("Boxplot of Medical Charges per Region")
```
Based from above plot, we can disclose that region of origin doesn’t have much impact with the amount of medical cost.

```{r}
# Smoking status
describeBy(assurance$charges,assurance$smoker)
```
```{r}
ggplot(data = assurance,aes(smoker,charges)) + geom_boxplot(fill = c(2:3)) +
  theme_classic() + ggtitle("Boxplot of Medical Charges by Smoking Status")
```
As we can see that the smokers spends a lot more in terms of medical expenses compared to non-smokers by almost 4x.

```{r}
#correlation
pairs.panels(assurance[c("age","sex", "bmi", "children","smoker", "charges")])
```



```{r}
cor(assurance[-7])[,"charges"] 

```

We can see that smoker and age has the highest correlation with charges amongst our numeric variable.

```{r}
#Model Building 
#Build a model with all variables
mod1 <- lm(charges ~ age + sex + bmi + children + smoker + region, data = assurance)
summary(mod1)

```
We utilize all the variables on this model, and the r-squared of 0.7509 is good, which implies that 75.09% of the variation of charges could be explained by the set of independent variables. we could also observe that all of the independent variables we have included with the exception of gender is a statistically significant predictor of medical charges (p-value less than 0.05 <- level of significance)

```{r}
#attach(assurance)
plot(assurance$age,assurance$charges,col=assurance$smoker)

```
```{r}
summary(assurance$charges[assurance$smoker=="no"])
```

```{r}
summary(assurance$charges[assurance$smoker=="yes"])
```

We could see a trend here, as people get older the higher their medical expenses would be which is kinda expected. But, regardless of age, smokers has higher medical expenses than non-smokers. I will try to create a model using only age and smoking status just for comparison. It really seems like smoker is the single most important variable in predicting medical charges.

```{r}

mod2<-lm(charges~age+smoker+bmi,data=assurance)
summary(mod2)

```

Using just age and smoker as independent variables, we have built a model with an r-squared of 72.14% which is comparable with our first model which use all of the original variables. In regression analysis, we would want to create a model that is accurate but at the same time as simple as possible. So if I would have to choose, I would select this second model than the first one. 



```{r}
qqnorm(studres(mod2))
qqline(stdres(mod2))
```

```{r}
plot(mod2)
```



```{r}
intercepts<-c(coef(mod2)["(Intercept)"],coef(mod2)["(Intercept)"]+coef(mod2)["smokeryes"])
lines.df<- data.frame(intercepts = intercepts,
                      slopes = rep(coef(mod2)["age"], 2),
                      smoker = levels(assurance$smoker))
qplot(x=age,y=charges,color=smoker,data=assurance)+geom_abline(aes(intercept=intercepts,slope=slopes,color=smoker),data=lines.df)  + scale_y_continuous(breaks = seq(0,65000,5000))
```
If we would visualize the most recent regression model we built, this is how it would like. As you can see, we have 2 parallel lines which indicates we have 2 different regression equations having the same slope but different intercepts. Slope of the regression lines is equal to the coefficient of the variable age (274.87). While in terms of the intercept, intercept for smokers is higher by 23,855.30 compared to non-smokers. This indicates that, on average, smokers’ medical charges is higher by about USD 24,000 adjusting for age. (Smoking is bad for your health!)



```{r}

```

```{r}

n_train <- round(0.8 * nrow(assurance))
train_indices <- sample(1:nrow(assurance), n_train)
Data_train <- assurance[train_indices, ]
Data_test <- assurance[-train_indices, ]

```
```{r}
#Saving R-squared
rsq0 <- summary(mod2)$r.squared


#predict data on test set
prediction0 <- predict(mod2, newdata = Data_test)
#calculating the residuals
residuals0 <- Data_test$charges - prediction_0
#calculating Root Mean Squared Error
rmse0 <- sqrt(mean(residuals_0^2))

print(paste0("RMSE for second model: ", round(rmse0, 2)))

```

```{r}
Data_test$prediction <- predict(mod2, newdata = Data_test)
ggplot(Data_test, aes(x = prediction, y = charges)) + 
  geom_point(color = "blue", alpha = 0.7) + 
  geom_abline(color = "red") +
  ggtitle("Prediction vs. Real values")
```
```{r}

ggplot(Data_test, aes(x = residuals0)) + 
  geom_histogram(bins = 15, fill = "blue") +
  ggtitle("Histogram of residuals")


```
```{r}

robert= data.frame(age = 19,
                  bmi = 27.9,
                  smoker = "yes")
print(paste0("Health care charges for robert: ", round(predict(mod2, robert), 2)))

```
```{r}
Lisa = data.frame(age = 40,
                   bmi = 50,
                   smoker = "no")
print(paste0("Health care charges for Lisa: ", round(predict(mod2, Lisa), 2)))
```

```{r}
julien =  data.frame(age = 30,
                   bmi = 31.2,
                   smoker = "no")
print(paste0("Health care charges for julien: ", round(predict(mod2, julien), 2)))
```

